Domain Specific Language for high performance computing

### 2024-02-09 paper/paper.tex
japanese of introduction ver1

\documentclass[ams, a4j]{U-AizuGT}
\usepackage{pifont}
% \usepackage{graphicx}
\usepackage{cite}
\usepackage{mathtools}
% \usepackage{listings,jvlisting}
% \documentclass[a4j]{jarticle} %ここは関係ない
\usepackage{listings,jvlisting} 
\usepackage[dvipdfmx]{graphicx}

@@

\bibliographystyle{ieice}
\author{Ryuki Hiwada}
\studentid{s1280076}
\supervisor{Naohito Nakasato}

@@

\begin{document}
\maketitle
\begin{abstract}

@@

\subsection{Background}
In astrophysics and astronomy, to numerically calculate the dynamical 
evolution of N particles interacting gravitationally, N-body simulations 
are required. Figure 1 shows the equation for interparticle interactions 
in N-body simulations. If the equation is naively computed, the time
complexity of calculation of interparticle interactions is 
\begin{math}O(N^2) \end{math}, where 
N is the number of particles. Therefore, parallelization is required to
speed up numerical simulations. To write a parallelized code for a 
numerical simulation, a user needs to understand the architecture of 
computer systems in  detail. If a parallelized code is automatically
generated by only describing the formulas and data of the numerical
simulation, the above problems are solved. To realize the parallelization,
we will develop Domain Specific Language  DSL \rparen, which is a 
programming language specialized to a domain, for example, SQL and HTML.
\subsection{parallelization}
SIMD is one of the categories in Flynn's taxonomy 1, related to computer arithmetic
 processing. Using SIMD, arithmetic operations can be applied simultaneously to
multiple data. In this paper, we compute and parallelize multiple particles. For example, 
as shown in Figure 1, the computation of acceleration for different particles is independent
of each other, allowing the process to be parallelized. As a result, for instance Assuming 
that the variables for particles are double-precision floating-point numbers, we can execute 
an operation on four elements simultaneously using the AVX2 instruction, it is possible to 
perform operations on four elements. Therefore, compared to non-parallelized code, there is
a potential to accelerate the computation by four times. We will describe the design of a 
DSL in the next section to generate such parallelized code from the formulation of particle-particle interactions.
\subsection{Overview of DSL}

We have created a Domain Specific Language  DSL \rparen named  Pyker \rparen. This language's 
processor can read code describing particle interaction formulas and generate code 
that accelerates these computations. The language allows for the definition of 
variables and the description of interaction formulas. In the variable definition 
section, it's possible to define variables such as the mass and position of 
particles, variables for storing results, and other necessary variables. For the formula
part, we have enabled calculations including basic arithmetic operations, sqrt, power 
functions, and computing the norm of vectors. From this description, a kernel function 
is generated, which can then be called and used. We have created a Domain Specific 
Language  DSL  named  Pyker. This language's processor can read code describing particle 
interaction formulas and generate code that accelerates these computations. The language allows
for the definition of variables and the description of interaction formulas. In the variable
definition section, it's possible to define variables such as the mass and position of 
particles, variables for storing results, and other necessary variables. For the formula part, 
we have enabled calculations including basic arithmetic operations, sqrt, power functions, and 
computing the norm of vectors. From this description, a kernel function is generated, which can then be called and used.
\subsection{Aim of this paper}

@@

\begin{lstlisting}[frame=single, caption=hoge]

#include<math>
int kernel( (1) double xi[][3], double xj[][3], double eps2, double ai[][3], ...) {
  //(2)
  int i;
  int j;
  double dr_v0;
  double dr_v1;
  double dr_v2;
  // ... 
  // ... 
  // ...

  for(i = 0;i < n;i += 1) {
    for(j = 0;j < n;j += 1) {
      //(3)
      dr_v0 = -xi[i][0] + xj[j][0];
      dr_v1 = -xi[i][1] + xj[j][1];
      dr_v2 = -xi[i][2] + xj[j][2];
      // ... 
      // ... 
      // ... 
    }
  }
}
\end{lstlisting}

@@

\begin{lstlisting}[frame=single, caption=Nbody-kernel.pikg, label=Nbody-kernel.pikg]
  EPI F64vec ri:r
  EPJ F64vec rj:r
  EPJ F64 mj:m

  EPJ F64 eps2:eps

  FORCE F64vec ai:acc

  rij = ri - rj
  r2 = rij * rij + eps2
  r_inv  = rsqrt(r2)
  r2_inv = r_inv * r_inv
  mr_inv  = mj * r_inv
  mr3_inv = r2_inv * mr_inv
  ai += mr3_inv * rij

@@

\section{Conclusion}
\section{Acknowledgement}
\bibliography{biblist}
\end{document}

>>>>

\documentclass[ams, a4j]{U-AizuGT}
\usepackage{pifont}
% \usepackage{graphicx}
\usepackage{cite}
\usepackage{mathtools}
% \usepackage{listings,jvlisting}
% \documentclass[a4j]{jarticle} %ここは関係ない
\usepackage{listings,jvlisting} 
\usepackage[dvipdfmx]{graphicx}
\usepackage{physics}
\bibliographystyle{ieicetr}

@@

% \bibliographystyle{ieice}
\author{Ryuki Hiwada}
\studentid{s1280076}
\supervisor{Naohito Nakasato}

@@

\begin{document}

\maketitle
\begin{abstract}

@@

\subsection{Background}

宇宙物理学や流体力学等、様々な分野で粒子系シミュレーションが行われている.
例えば、惑星間の重力による力学的進化をシミュレーションしたN体シミュレーションや
水の流体運動のモデル化がある。
これらのシミュレーションには粒子の数に対して計算量が膨大になる。
例としてN体シミュレーションの場合を説明する。N体シミュレーションでは,
ほとんどの計算時間を重力相互作用の計算が占めている。\eqref{eq:gravity} は、
重力相互作用の式である.
\begin{equation}
a_i = \sum_{\substack{j \neq i}}^{N} G m_j \frac{\vb{x}_j - \vb{x}_i}{\left( \left| \vb{x}_j - \vb{x}_i \right| + \epsilon^2 \right)^{3/2}\label{eq:gravity}}
\end{equation}

ここで $a_i$, $\vb{x}_i$はそれぞれ粒子iの加速度と位置を表し、
$m_j$,$\vb{x}_j$はそれぞれ粒子jの質量と位置を表し、
$N$,$G$,$\epsilon$はそれぞれ粒子の数、重力定数発散を防ぐためのソフトニングパラメータである。
\eqref{eq:gravity}から計算量は粒子数Nに比例して、\begin{math}O(N^2) \end{math}となる。
N体シミュレーションでは、粒子の数が膨大になる場合が多い。例えば、
球状星団という天体には星が約$10^4 ~ 10^5 $ 個存在する。このため非常に長い計算時間がかかるため高速化が求めれる。
粒子間相互作用の計算は、粒子同士の計算が独立しているためデータ並列性が高い。そのため、単一の命令に対して複数の
データを処理するSIMDによる並列に適している。
SIMDによる並列計算の例として、小池らが行ったGRAPE-DR\cite{GRAPE-explain-thesis}による重力多体問題シミュレーションがある。\cite{GRAPE-DR-thesis}
しかし、SIMD計算を行うには,コンピュータのアーキテクチャを考慮したコードを記述しなくてはならない。

この研究の目的は、ユーザーが粒子間相互作用の計算を並列化を意識することなく、並列に実行できるように
することである。そのためにPykerというDomain Specific Language\lparen DSL \rparen を開発した。
Pykerでは、ユーザに必要な変数の定義と数式から相互作用の計算をSIMD命令によって並列化したコードを生成する。
Pykerの評価方法としては、C言語のコンパイラが自動で行うSIMD化とPykerによるSIMD化の性能を比較する。
そのために、3つの相互作用の計算を行うコードをそれぞれにPykerから生成されたSIMD命令を適応しない
場合のコードと、SIMD命令を適応したコードの実行時間を計測し比較する。
また、Pykerを作成するにあたって粒子間相互作用の計算するコードを生成するPIKGを参考にした。

@@

@@

\begin{lstlisting}[frame=single, caption=Nbody-kernel.pikg, label=Nbody-kernel.pikg]
  EPI F64vec ri:r
  EPJ F64vec rj:r
  EPJ F64 mj:m
  EPJ F64 eps2:eps
  FORCE F64vec ai:acc
  rij = ri - rj
  r2 = rij * rij + eps2
  r_inv  = rsqrt(r2)
  r2_inv = r_inv * r_inv
  mr_inv  = mj * r_inv
  mr3_inv = r2_inv * mr_inv
  ai += mr3_inv * rij

@@

\section{Conclusion}
\section{Acknowledgement}

\begin{thebibliography}{99}
\end{thebibliography}
\bibliography{biblist}

\end{document}
